Very simplified likelihood

### 2024-02-13 inst/very_simplified_likelihood.Rmd
simplifications and corrections

\[
\frac{(B-1)\Gamma(B+n-r-1)}{\Gamma(B+n)}
\]

@@

\[
\frac{1}{(B+3)(B+2)(B+1)B}=\frac{(B-1)!}{(B+3)!}=\frac{\Gamma(B)}{\Gamma(B+4)}
\]

@@

\[
\frac{1}{(B+n-1)(B+n-2)\cdots(B+1)B}=\frac{(B-1)!}{(B+n)!}=\frac{\Gamma(B)}{\Gamma(B+n+1)}
\]

@@

```{r}
f_single <- function(a,r,n,log=FALSE){  # not vectorised
  B <- 1/(1-a)
  if(log){
     if(r==n){  # last place
       out <- lgamma(B)-lgamma(B+n)
     } else {   # not last
       out <- log(B-1) - sum(log(B + (n-r-1):(n-1)))
     }
   } else {  # not-log 
     if(r==n){
      out <- gamma(B)/gamma(B+n)
     } else {
       out <- (B-1)/prod(B+(n-r-1):(n-1))
     }
   }	
   return(out)
}

f_single2 <- function(a,r,n,log=FALSE){  # not vectorised
  B <- 1/(1-a)
  if(log){
     if(r==n){  # last place
       out <- lgamma(B)-lgamma(B+n)
     } else {   # not last
       out <- log(B-1) - lfactorial(B+n-1) + lfactorial(B+n-r-2)
     }
   } else {  # not-log 
     if(r==n){
      out <- gamma(B)/gamma(B+n)
     } else {
       out <- (B-1)/prod(B+(n-r-1):(n-1))
     }
   }	
   return(out)
}

@@

  if(log){
    out <- 0
    for(i in seq_len(nrow(M))){out <- out + f_vec_a(a,r=M[i,1],n=M[i,2],log=TRUE)}
  } else {
    out <- 1
    for(i in seq_len(nrow(M))){out <- out * f_vec_a(a,r=M[i,1],n=M[i,2],log=FALSE)}
  }
  return(out)
}

@@

```{r}
a <- 0.23423434
b <- 1-a
B <- 1/b
n <- 9
r <- 3
c(
b^3*a/((a+9*b) * (a+8*b) * (a+7*b) * (a+6*b)),
(B-1)/prod(B+(5:8)),
f_vec_arn(a,r,n,log=FALSE))
```

```{r}
out <- list()
for(n in 1:10){
  out[[n]]  <- rep(NA,n+1)
  for(r in 0:n){
    if(r==0){
      jj <- 1
    } else if(r==n){n
      jj <- 0
    } else {
      jj <- MLE(r,n)$maximum
    }  
    out[[n]][r+1] <- jj
  }
} 
out
```

```{r}
plotterp <- function(...){
plot(NA,xlim=c(1,10),ylim=c(0,1),type="n",xlab="n",ylab=expression(hat(a)))
for(n in 1:10){
  for(r in 0:n){   points(n,out[[n]][r+1],pch=16) }
  for(r in 0:n){   text(n+0.2,out[[n]][r+1],r,cex=0.5,col='gray') }
}
}
plotterp()
pdf(file="dotprobs.pdf")
plotterp()
dev.off()
```

@@

```{r calculatelikes,cache=F}
a <- seq(from=0.15,to=0.5,by=0.01)
La <- f_vec_arn(a,r=results$place-1,results$runners-1,log=TRUE)
La <- La - max(La,na.rm=TRUE)
```

```{r plotlike}
plot(a,La)
abline(h=c(0,-2))
```

@@

```{r calculatelikes1,cache=F}
a <- seq(from=0.005,to=0.6,by=0.005)
La <- f_vec_arn(a,r=results$place[1]-1,results$runners[1]-1,log=TRUE)
La <- La - max(La,na.rm=TRUE)
plot(a,La,type='b')
abline(h=c(0,-2))
```

@@

```{r, label=onedimopt}
f <- function(a){supp2(c(a,a),results$place,results$runners)}
a <- seq(from=0.1,to=0.45,by=0.01)
alike <- sapply(a,f)
plot(a,alike-max(alike,na.rm=TRUE),type='b')
abline(h=c(0,-2))
optimize(f,c(0.1,0.6),maximum=TRUE)
```

@@

```{r}
a <- seq(from=0.1,by=0.01,to=0.7)
wp <- category=='pure'
wa <- category=='applied'
```

@@

```{r}
Lpure <- f_vec_arn(a,rank[wp]-1,class_size[wp]-1,log=TRUE)
Lpure <- Lpure - max(Lpure,na.rm=TRUE)
plot(a,Lpure)
optimize(function(a){f_vec_arn(a,rank[wp]-1,class_size[wp]-1,log=TRUE)},c(0.3,0.6),maximum=TRUE)
optimize(function(a){f_vec_arn(a,rank[wa]-1,class_size[wa]-1,log=TRUE)},c(0.3,0.6),maximum=TRUE)
```

>>>>

\[
\frac{(B-1)(B+n-r-2)!}{(B+n-1)!}
\]

@@

\[
\frac{1}{(B+3)(B+2)(B+1)B}=\frac{(B-1)!}{(B+3)!}
\]

@@

\[
\frac{1}{(B+n-1)(B+n-2)\cdots(B+1)B}=
\frac{(B-1)!}{(B+n-1)!}
\]

@@

```{r,label=definefsingle}
f_single_old <- function(a,r,n,log=FALSE){  # not vectorised
  B <- 1/(1-a)
  if(log){
     if(r==n){  # last place
       out <- lgamma(B)-lgamma(B+n)
     } else {   # not last
       out <- log(B-1) - sum(log(B + (n-r-1):(n-1)))
     }
   } else {  # not-log 
     if(r==n){
      out <- gamma(B)/gamma(B+n)
     } else {
       out <- (B-1)/prod(B+(n-r-1):(n-1))
     }
   }	
   return(out)
}

f_single_old2 <- function(a,r,n,log=FALSE){  # not vectorised
  B <- 1/(1-a)
  if(log){
     if(r==n){  # last place
       out <- lgamma(B)-lgamma(B+n)
     } else {   # not last
      out <- log(B-1) - lgamma(B+n) + lgamma(B+n-r-1)
     }
   } else {  # not-log 
     if(r==n){
      out <- gamma(B)/gamma(B+n)
     } else {
       out <- (B-1)/prod(B+(n-r-1):(n-1))
     }
   }	
   return(out)
}

f_single <- function(a,r,n,log=FALSE){  # not vectorised
  B <- 1/(1-a)
  if(log){
      out <- log(B-1) - lgamma(B+n) + lgamma(B+n-r-1)
   } else {  # not-log 
      out <- (B-1)/prod(B+(n-r-1):(n-1))
   }	
   return(out)
}

@@

  if(log){
    out <- 0
    for(i in seq_len(nrow(M))){out <- out + f_vec_a(a,r=M[i,1],n=M[i,2],log=TRUE)}
  } else {
    out <- 1
    for(i in seq_len(nrow(M))){out <- out * f_vec_a(a,r=M[i,1],n=M[i,2],log=FALSE)}
  }
  return(out)
}
```

```{r,label=defmle}

@@

```{r,label=verifytwo}
a <- 0.23423434
b <- 1-a
B <- 1/b
n <- 9
r <- 3
c(
way1 = b^3*a/((a+9*b) * (a+8*b) * (a+7*b) * (a+6*b)),
way2 = (B-1)/prod(B+(5:8)),
way3 = f_vec_arn(a,r,n,log=FALSE)
)
```

```{r,label=makejj}
out <- list()
for(n in 1:10){
  out[[n]]  <- rep(NA,n+1)
  for(r in 0:n){
    if(r==0){
      jj <- 1
    } else if(r==n){n
      jj <- 0
    } else {
      jj <- MLE(r,n)$maximum
    }  
    out[[n]][r+1] <- jj
  }
} 
out
```

```{r,label=plotjj}
plotterp <- function(...){
plot(NA,xlim=c(1,10),ylim=c(0,1),type="n",xlab="n",ylab=expression(hat(a)))
for(n in 1:10){
  for(r in 0:n){   points(n,out[[n]][r+1],pch=16) }
  for(r in 0:n){   text(n+0.2,out[[n]][r+1],r,cex=0.5,col='gray') }
}
}
plotterp()
pdf(file="dotprobs.pdf")
plotterp()
dev.off()
```

@@

```{r calculatelikes,cache=F}
a <- seq(from=0.25,to=0.65,by=0.01)
La <- f_vec_arn(a,r=results$place-1,results$runners-1,log=TRUE)
La <- La - max(La,na.rm=TRUE)
```

```{r plotlike}
plot(a,La,ylab='support')
abline(h=c(0,-2))
```

@@

```{r calculatelikes1,cache=F}
a <- seq(from=0.005,to=0.75,by=0.01)
La <- f_vec_arn(a,r=results$place[1]-1,results$runners[1]-1,log=TRUE)
La <- La - max(La,na.rm=TRUE)
plot(a,La,type='b',ylab='fishy')
abline(h=c(0,-2))
```

@@

```{r, label=onedimopt}
f <- function(a){supp2(c(a,a),results$place,results$runners)}
a <- seq(from=0.2,to=0.6,by=0.01)
alike <- sapply(a,f)
plot(a,alike-max(alike,na.rm=TRUE),type='b',ylab='blong')
abline(h=c(0,-2))
optimize(f,c(0.1,0.6),maximum=TRUE)
```

@@

```{r}
a <- seq(from=0.2,by=0.01,to=0.8)
wp <- category=='pure'
wa <- category=='applied'
```

@@

```{r}
Lpure <- f_vec_arn(a,rank[wp]-1,class_size[wp]-1,log=TRUE)
Lpure <- Lpure - max(Lpure,na.rm=TRUE)
plot(a,Lpure,ylab='pure strength')
optimize(function(a){f_vec_arn(a,rank[wp]-1,class_size[wp]-1,log=TRUE)},c(0.3,0.6),maximum=TRUE)
optimize(function(a){f_vec_arn(a,rank[wa]-1,class_size[wa]-1,log=TRUE)},c(0.3,0.6),maximum=TRUE)
```
